Automatic Matching of Bullet Lands

### 2015-12-29 imaging-paper.Rnw
cleaned up probability ratios, included ROC curves

qplot(x, value, group=variable, geom="polygon", data=dm, fill=variable, 
      alpha=I(0.6), colour=I("grey20")) + 
  facet_wrap(~var, ncol=4, scales="free") +
  theme_bw() + ylab("density") + xlab("") +
  scale_fill_brewer("Matches", palette="Paired") +
  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
        legend.background = element_rect(colour="grey75"),
        plot.margin = unit(c(0,0,0,0), unit="cm"))
@
\caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}.}
\end{figure}

<<prob.ratios, echo=FALSE>>=
library(reshape2)
dens.stats <- dm %>% group_by(var, variable) %>% mutate(
  sum = sum(value),
  up = cumsum(value)/sum,
  down = 1 - cumsum(value)/sum
)
ratios <- dens.stats %>% group_by(var, x) %>% summarise(
  upper = sum(up[variable=="KM"])/sum(up[variable=="KNM"]),
  lower = sum(down[variable=="KM"])/sum(down[variable=="KNM"])
)
rm <- melt(ratios, measure.var=c("upper", "lower"))
levels(rm$variable) <- c("P(X < c | KM) / P(X < c | KNM)", "P(X > c | KM) / P(X > c | KNM)")
qplot(x, value, colour=variable,  data=rm) + facet_wrap(~var, scales="free") + scale_y_log10() + theme_bw() + geom_hline(yintercept=1) + scale_colour_brewer(palette="Set2") + ylab("Probability Ratios")
@

>>>>

qplot(x, value, group=variable, geom="polygon", data=dm, fill=variable, 
      alpha=I(0.6), colour=I("grey20")) + 
  facet_wrap(~var, ncol=4, scales="free") +
  theme_bw() + ylab("density") + xlab("") +
  scale_fill_brewer("Matches", palette="Paired") +
  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
        legend.background = element_rect(colour="grey75"),
        plot.margin = unit(c(0,0,0,0), unit="cm"))
@
\caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}. Shifts in the mode of the density functions between known matches and known non-matches indicate the variable's predictive power in distinguishing matches and non-matches. }
\end{figure}

\begin{figure}[hbtp]
  \centering
<<ratios-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
library(reshape2)
dens.stats <- dm %>% group_by(var, variable) %>% mutate(
  sum = sum(value, na.rm=TRUE),
  up = cumsum(value)/sum,
  down = 1 - cumsum(value)/sum
)
levels(dens.stats$variable) <- c("KM", "KNM")
ratios <- dens.stats %>% group_by(var, x) %>% summarise(
  lower = sum(up[variable=="KM"], na.rm=TRUE)/sum(up[variable=="KNM"], na.rm=TRUE),
  upper = sum(down[variable=="KM"], na.rm=TRUE)/sum(down[variable=="KNM"], na.rm=TRUE)
)
rm <- melt(ratios, measure.var=c("lower", "upper"))
rm$value <- pmin(pmax(rm$value, 0.1), 10)
levels(rm$variable) <- c("P(X < c | KM) / P(X < c | KNM)", "P(X > c | KM) / P(X > c | KNM)")
qplot(x, value, colour=variable,  data=rm, geom="line", group=variable, 
      size=I(1)) + xlab("X") +
  facet_wrap(~var, scales="free", ncol=4) + 
  scale_y_log10(limits=c(1/10,10)) + theme_bw() + 
  geom_hline(yintercept=1, size=0.5) + 
  scale_colour_brewer("Estimated Ratios", palette="Set2") + 
  ylab("(Logged) Probability Ratios") +
  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
        legend.background = element_rect(colour="grey75"),
        plot.margin = unit(c(0,0,0,0), unit="cm")) 
@
\caption{\label{fig:ratios}Estimated probability ratios for all of the features described in section~\ref{sec:algorithm}. Ratios are logged and and artificially constrained to an interval between (1/10 and 10). Large deviations from one, indicate a strong distinction between matches and known matches based on the variable. Both left tail and right tail probabilities are shown. }
\end{figure}

\begin{figure}[hbtp]
  \centering
<<rocs-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
# plot false positive against false negative
# prob of false positive: P(match | KNM)
# prob of false negative: P(non-match | KM)

# for  X < c: we get probability for P(X < c | KNM) and P(X < c | KM)

features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
errors <- plyr::ldply(features, function(x) {
  xx <- unique(bstats[,x])
  if (length(xx) > 500) {
    xr <- range(bstats[,x])
    xx <- seq(xr[1], xr[2], length.out=500)
  }
  # upper and lower rule:
  # upper: match is defined as X > xx
  # lower: match is defined as X < xx
  subbstats <- subset(bstats, flagged == FALSE)
  subbstats$match <- factor(subbstats$match)

  errors <- plyr::ldply(xx, function(cc) {
    idx <- which(subbstats[,x] >= cc)
    upper <- data.frame(xtabs(~match, data= subbstats[idx,,drop=FALSE])/
                          xtabs(~match, data= subbstats))
    idx <- which(subbstats[,x] <= cc)
    lower <- data.frame(xtabs(~match, data= subbstats[idx,,drop=FALSE])/
                          xtabs(~match, data= subbstats))

    data.frame(value = cc, match=lower[,1], lower=lower[,2], upper=upper[,2])
  })

  errors$variable <- x
  errors
})

rocs.lower <- dcast(errors, variable+value~match, value.var="lower")
rocs.lower$type <- "lower"
rocs.upper <- dcast(errors, variable+value~match, value.var="upper")
rocs.upper$type <- "upper"
rocs <- rbind(rocs.lower, rocs.upper)

rocs$variable <- factor(rocs$variable, levels=features)
qplot(`FALSE`, `TRUE`, data=rocs,  geom="line", group=type) + theme_bw() + geom_abline(linetype=2) + xlab("False negative rate") + 
  ylab("False positive rate") + facet_wrap(facets=~variable, ncol=4)
@
\caption{\label{fig:rocs}ROC curves for all of the features described in section~\ref{sec:algorithm}. \hh{XXX compare ROC curve of CMS against the other individual features -- ccf should be a lot more predictive than CMS. XXX}
\hh{XXX next steps: get rid of lower curve. find area under the curve, put value on bottom right of the chart, sort by ROC.}}
\end{figure}
